Learning to Infer Graphics Programs from Hand-Drawn Images

### 2018-04-25 nips_2018.tex
nips 2018

% if you need to pass options to natbib, use, e.g.:
\PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2017
%
% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2017}

%\usepackage{nips_2017}
\usepackage{dsfont}
% to compile a camera-ready version, add the [final] option, e.g.:
%\usepackage[final]{nips_2017}
%ICLR packages:
\usepackage{iclr2018_conference,times}
%\iclrfinalcopy

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{listings}
\lstset{basicstyle = \small\ttfamily}
\usepackage{graphicx} % more modern
%\usepackage{epsfig} % less modern
%\usepackage{subfig} 
\usepackage{fancyvrb}

@@

%\lhead{nothing}

@@

\begin{abstract}
  We introduce a model that learns to convert simple hand drawings
  into graphics programs written in a subset of \LaTeX.~The model
  combines techniques from deep learning and program synthesis.  We
  learn a convolutional neural network that proposes plausible drawing
  primitives that explain an image. These drawing primitives are
  like a trace of the set of primitive commands issued by a graphics
  program. We learn a model that uses program synthesis techniques to
  recover a graphics program from that trace. These programs have constructs like variable
  bindings, iterative loops, or simple kinds of conditionals. With a
  graphics program in hand, we can correct errors made by the deep
  network %, measure similarity between drawings by use of similar high-level geometric
  %structures,
  and extrapolate drawings.  Taken together these results
  are a step towards agents that induce useful, human-readable
  programs from perceptual input.
\end{abstract}

@@

 How can an agent convert noisy, high-dimensional perceptual input
 to a symbolic, abstract object, such as a computer program?  Here we
 consider this problem within a graphics program synthesis domain.  We
 develop an approach for converting  hand
 drawings into executable source code for drawing the original image.
 The graphics programs in our domain draw simple figures like those found in
 machine learning papers (see Fig.~\ref{firstPageExamples}a).
 \begin{figure}[H]\vspace{-0.7cm}
  \begin{minipage}[t]{0.7\linewidth}  
\begin{tabular}{llll}
  \includegraphics[width = \exampleImageSize]{figures/expert-60.png}&
  \includegraphics[width = \exampleImageSize]{figures/expert-5.png}&
    \includegraphics[width = \exampleImageSize]{figures/expert-17.png}&
    \includegraphics[width = \exampleImageSize]{figures/expert-58.png}\\
  \includegraphics[width = \exampleImageSize]{figures/60.png}&
  \includegraphics[width = \exampleImageSize]{figures/5.png}&
    \includegraphics[width = \exampleImageSize]{figures/17.png}&
    \includegraphics[width = \exampleImageSize]{figures/58.png}
\end{tabular}
\subcaption{}
  \end{minipage}%
  \begin{minipage}[t]{0.3\linewidth}
    \begin{tikzpicture}
      \node(picture) at (0,0) {\includegraphics[width = 2cm]{figures/expert-31.png}};
      \draw[very thick,->] (picture.south)  -- (0,-2);
    \end{tikzpicture}\\
    \begin{minipage}[t]{0.3\linewidth}
\begin{Verbatim}
for (i < 3)
 rectangle(3*i,-2*i+4,
           3*i+2,6)
 for (j < i + 1)
  circle(3*i+1,-2*j+5)
\end{Verbatim}      
      \end{minipage}
    \subcaption{}
  \end{minipage}
  \caption{(a): Model learns to convert hand drawings (top) into \LaTeX~(rendered below). (b) Synthesizes high-level \emph{graphics program} from hand drawing.}\label{firstPageExamples}
\end{figure}

The key observation behind our work is that generating a programmatic representation from an image of a diagram  involves two distinct steps that require different technical approaches. The first step involves identifying the components such as rectangles, lines and arrows that make up the image. The second step involves identifying the high-level structure in how the components were drawn. In Fig. 1(b), it means identifying a pattern in how the circles and rectangles are being drawn that is best described with two nested loops, and which can easily be extrapolated to a bigger diagram.

We present a hybrid architecture for inferring graphics programs that is structured around these two steps. For the first step,   a deep network to infers a set of primitive shape-drawing commands. We refer to this set as a \emph{trace set} since it corresponds to the set of commands in a program's execution trace but lacks the high-level structure that determines how the program decided to issue them.
%For added robustness, we train the network to produce a proposal distribution of the most likely traces and use stochastic search to find the trace that best matches the image.
%The network is trained from an automatically-generated corpus of synthetic image-generating programs.
The second step involves synthesizing a high-level program capable of producing the trace set identified by the first phase.
%In principle, we could also train a neural network to learn this mapping. In practice, though, this would be challenging because it would no longer be sufficient to train the network with a synthetic corpus of randomly generated programs, since our goal is to generate programs that satisfy certain semantic constraints that ensure that we get the program that the user most likely intended.
%This paper argues that this stage is best
We achieve this by \emph{constraint-based program synthesis} \citep{solar2008program}. The program synthesizer searches the space of possible programs for one capable of producing the desired trace set -- inducing structures like
symmetries, loops, or conditionals.
Although these program synthesizers do not need any training data,
we show how to learn a \emph{search policy} 
in order to synthesize programs an order of magnitude faster than constraint-based synthesis techniques alone.

One might be tempted to try to go directly from perceptual input (an image) to a program,
much like recent neural models of program induction that regress from problem statements to programs (e.g.,~\cite{devlin2017robustfill}). We advocate an alternative framing, which we call The Trace Hypothesis:

%% Images are high-dimensional and unstructured while programs are high-level
%% and symbolic. We resolve this mismatch by advancing a new model of
%% collaboration between these two representations, which we call The Trace Hypothesis:

\noindent \textbf{The Trace Hypothesis.} The set of commands issued by a program,
which we call a \emph{trace set},
is the correct liaison between high-dimensional unstructured perceptual input and high-level structured symbolic representations.

The roadmap of our paper is structured around the trace hypothesis, as outlined in Fig.~\ref{roadmap}.
\begin{figure}[h]
  \centering\begin{tikzpicture}
    \node[ thick,anchor = west,inner sep=0pt,label=below:{\small Perceptual input}](observation) at (0,0) {\includegraphics[width = 1.5cm]{figures/expert-39-trimmed.png}};
    \node[ultra thick,anchor = west,inner sep=0pt](traceSource) at (3,0.5){    \begin{lstlisting}[basicstyle = \scriptsize\ttfamily]
line, line,
rectangle,
line, ...
\end{lstlisting}};
    \node[ultra thick,anchor = west,inner sep=0pt](traceImage) at (3.3,-0.5) {
      \includegraphics[width = 0.9cm]{figures/39-parse.png}}; 
    \node(trace)[draw,thin,fit = (traceImage) (traceSource), label = below:{{\small Trace set}}] {};

    \node[draw, thick,anchor = west,inner sep=2pt,label=below:{\small Program}](program) at (6.3,0) {
      \begin{lstlisting}
for (j < 3)
for (i < 3)
if (...)
 line(...)
 line(...)
rectangle(...)
    \end{lstlisting}};
    \node[ultra thick,anchor = west,inner sep=0pt,label=below:{\small Extrapolation}](extrapolate) at (9.9,-0.4) {\includegraphics[width = 1.3cm]{figures/39-extrapolated.png}};
%    \node[ultra thick,anchor = west,inner sep=0pt,label=below:{\small Similarity}](similarity) at (11.3,0) {\includegraphics[width = 1cm]{figures/expert-38-trim.png}};
    \node[ultra thick,anchor = west,text width = 1.3cm,inner sep=0pt](errors) at (10,1) {\small Error correction};

    \draw[->, thick] (observation.east) -- node[draw,fill = white,rotate = -90] {{\small Neural network}} (trace.west);
    \draw[->, thick] (trace.east) -- node[draw,fill = white,rotate = -90] {{\small Program synthesis}} (program.west);
    \draw[->, thick] (program.east) to[out = 80,in = 180] (errors.west);
%    \draw[->, thick] (program.east) to[out = 40,in = -230] (similarity.west);
        \draw[->, thick] (program.east) to[out = -90,in = 180] (extrapolate.west);

    \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
    ([yshift = -15,xshift = -115]trace.south) -- node[below = 6pt] {{\small Section~\ref{neuralNetworkSection}: Perception$\to$Trace set}} ([yshift = -15,xshift = -5]trace.south);
    \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
    ([xshift = 5,yshift = -15]trace.south) -- node[below = 6pt] {{\small Section~\ref{programSynthesisSection}: Trace Set$\to$Program}} ([xshift = 140,yshift = -15]trace.south);
    \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
([xshift = 150,yshift = -15]trace.south) -- node[below = 6pt] {{\small Section~\ref{applicationsSection}: Applications}} ([xshift = 230,yshift = -15]trace.south);
  \end{tikzpicture}
  \caption{Both the paper and the system pipeline are structured around the \emph{trace hypothesis}}\label{roadmap}
\end{figure}

The new contributions of this work are:
(1) The trace hypothesis: a framework for going from perception to programs,
which connects this work to other trace--based models,
like the  Neural Program Interpreter~\citep{DBLP:journals/corr/ReedF15};
(2) A model based on the trace hypothesis that converts sketches to high-level programs:
in contrast to converting images to vectors or low-level parses~\citep{huang2017shape,Nishida:2016:ISU:2897824.2925951,nsd,DBLP:journals/corr/Beltramelli17,im2latex}.
(3) A generic algorithm for learning a policy for efficiently searching for programs,
building on Levin search~\citep{levin1973universal}
and recent work like DeepCoder~\citep{BalGauBroetal16}.

Even with the high-level idea of a trace set, going from hand drawings to programs remains difficult. We address  these challenges:
(1) Inferring trace sets from images requires \emph{domain-specific} design choices from the deep learning and computer vision toolkits (Sec.~\ref{neuralNetworkSection}). (2) Generalizing to noisy hand drawings requires learning a \emph{domain-specific} noise model that is invariant to the variations across hand drawings (Sec.~\ref{generalizingTheHandDrawings}).  (3) Discovering good programs requires solving a difficult combinatorial search problem, because the programs are often long and complicated (e.g., 9 lines of code, with nested loops and conditionals). We give a \emph{domain-general} framework for learning a search policy that quickly guides program synthesizers toward the target programs (Sec.~\ref{learningASearchPolicy}).

%under a quantitative objective that maximizes the likelihood according to a prior distribution that aims to capture the user’s preference.

%% The key observation behind our work is that generating a programmatic representation from an image of a diagram actually involves two distinct steps that require different technical approaches. The first step involves identifying the components such as boxes, lines and arrows that make up the image. The second step involves identifying the high-level structure in how the components were drawn; for example, in Fig. 1(b), it means identifying that there is a pattern in how the circles and rectangles are being drawn that is best described with two nested loops, and which can easily be extrapolated to a bigger diagram.

%% We present a hybrid architecture for inferring graphics programs that is structured around these two steps. For the first step, our approach uses a deep neural network to infer a sequence of primitive shape-drawing commands that can generate an image similar to the observed image. We refer to this sequence as an \emph{execution trace}, since it corresponds to the sequence of primitive commands that the desired program would have issued, but it lacks the high-level structure that determines how the program decided to issue those commands. For added robustness, we train the network to produce a proposal distribution of the most likely traces, and use stochastic search to find the trace that best matches the input. The network is trained from an automatically generated corpus of synthetic image generating programs.

%% High dimensional perceptual input may seem ill matched to the abstract
%% semantics of a programming language. But programs with constructs such as
%% recursion or iteration produce a simpler \emph{execution trace} of
%% primitive actions; for our domain, the primitive actions are drawing
%% commands. Our hypothesis is that the execution trace of the program is
%% better aligned with the perceptual input, and that the trace can act
%% as a kind of bridge between perception and programs. We test this
%% hypothesis by developing a model that learns to map from an image to
%% the execution trace of the graphics program that drew it.  With the
%% execution trace in hand, we can bring to bear techniques from the
%% program synthesis community to recover the latent graphics program.
%% This family of techniques, called \emph{constraint-based program synthesis}~\cite{solar2008program},
%% work by modeling a set of possible programs inside of a constraint solver,
%% such as a SAT or SMT solver~\cite{de2008z3}.
%% These techniques excel at uncovering high-level symbolic structure,
%% but are not well equipped to deal with real-valued perceptual inputs.

%% We develop a hybrid architecture for inferring graphics programs.  Our
%% approach uses a deep neural network infer an execution trace from an
%% image; this network recovers primitive drawing operations such as
%% lines, circles, or arrows, along with their parameters. For added
%% robustness, we use the deep network as a proposal distribution for a
%% stochastic search over execution traces.  Section~\ref{neuralNetworkSection} describes this first stage of the architecture where we infer drawing commands from images, and explains how we handle noisy hand drawings.
%% Finally, we use program synthesis techniques to recover the program from its
%% trace.  The program synthesizer discovers constructs such as loops and
%% geometric operations such as reflections and affine transformations.
%% % \remark{This paragraph is all about making things a bit more
%% Section~\ref{programSynthesisSection} describes how the architecture
%% synthesizes programs from execution traces,
%% and how those programs are used for  correcting errors made by the deep network, measuring similarity between hand drawings,
%% and extrapolating figures.

%   so you really need more specifics about program synth here.}

%% Each of these three components -- the deep network, the stochastic
%% search, the program synthesizer -- confers its own advantages. From
%% the deep network, we get a fast system that can recover plausible
%% execution traces in about a minute~\remark{A minute seems slow to me, for deep net inference...}. From the stochastic search, we get
%% added robustness: essentially, the stochastic search can correct
%% mistakes made by the deep network's proposals.  From the program
%% synthesizer, we get abstraction: our system recovers coordinate
%% transformations, for loops, and subroutines, which are useful for
%% downstream tasks and can help correct some mistakes of the earlier stages.
%% % \remark{I wonder if this would work even better as a bulleted list...}
%% \remark{Tie these claims into the paper results: state what the `downstream' tasks are that you actually do, and refer to the sections where they occur}

\section{Neural architecture for inferring trace sets}\label{neuralNetworkSection}

We developed a deep network architecture for efficiently inferring a
trace set, $T$, from an image, $I$.
Our model combines ideas from
Neurally-Guided Procedural Models~\citep{ritchie2016neurally}
and Attend-Infer-Repeat~\citep{eslami1603attend}.
The network constructs the
trace set one drawing command at a time, conditioned on what it has drawn so far.
%When predicting the next drawing
%command, the network takes as input the target image $I$ as well as
%the rendered output of previous drawing commands.
%Intuitively, the network looks at the image it wants to explain, as well as what it has already drawn.  It then decides either to stop drawing or proposes another drawing command to add to the trace set; if it decides to continue drawing, the predicted primitive is rendered to its ``canvas'' and the process repeats.
Fig.~\ref{architecture} illustrates this architecture.  We first
pass a $256\times 256$ target image and a rendering of the trace set so
far (encoded as a two-channel image) to a convolutional network. Given
the features extracted by the convnet, a multilayer perceptron then
predicts a distribution over the next drawing command to add to the
trace set (see Tbl.~\ref{drawingCommandTable}).
%We predict the drawing command token-by-token,
%\footnote{Tokens are atomic symbols in the programming language  -- so the model is \emph{not} predicting lines of code character-by-character, but instead the model knows about the syntactic structure of primitive drawing commands}
%%   conditioning each token both on the image
%% features and on the previously generated tokens. 
We also use a
differentiable attention mechanism (Spatial Transformer
Networks:~\cite{jaderberg2015spatial}) to let the model attend to
different regions of the image while predicting drawing commands.
%In Fig.~\ref{architecture} the network first decided to emit the \verb|circle| token
%conditioned on the image features, then it emitted the $x$ coordinate of
%the circle conditioned on the image features and the \verb|circle|
%token, and finally it predicted the $y$ coordinate of the circle
%conditioned on the image features, the \verb|circle| token, and the
%$x$ coordinate.
We currently constrain
coordinates to lie on a discrete $16\times 16$ grid,
but the grid could be made arbitrarily fine.  %% See supplement for the full details
%% of the architecture, which we implemented in
%% Tensorflow~\citep{tensorflow2015-whitepaper}.
%% \remark{There are some more details that are important to provide
%%   about this architecture in the supplement: the
%%   functional form(s) of the probability distributions over tokens, the
%%   network layer sizes, which MLPs share parameters, etc.}

@@

  \node[draw,ultra thick,anchor = west,inner sep=0pt,minimum width = 2cm,minimum height = 3cm] (CNN) at (4.5,-2.5) {CNN};
  \node[inner sep = 0pt](tensorProduct) at ([xshift = -1.5cm]CNN.west) {$\bigotimes$};

@@

%  \node at (canvas.x,canvas.y) {Canvas};
\end{tikzpicture}
\caption{Our neural architecture for inferring the trace set of a graphics program from its output. \textcolor{blue}{Blue}: network inputs. Black: network operations. \textcolor{red}{Red}: samples from a multinomial. \texttt{Typewriter font}: network outputs. Renders snapped to a $16\times 16$ grid, illustrated in \textcolor{gray}{gray}. STN (spatial transformer network) is a differentiable attention mechanism~\citep{jaderberg2015spatial}.}  \label{architecture}
\end{figure}
\begin{table}[t]
  \caption{Primitive drawing commands currently supported by our model.}
\label{drawingCommandTable}
\begin{tabular}{ll}\toprule
  \begin{tabular}{l}
    \verb|circle|$(x,y)$
  \end{tabular}& \begin{tabular}{l}
    Circle at $(x,y)$
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|rectangle|$(x_1,y_1,x_2,y_2)$
  \end{tabular}&\begin{tabular}{l}
    Rectangle with corners at $(x_1,y_1)$ \& $(x_2,y_2)$
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|line|$(x_1,y_1,x_2,y_2,$\\
    \hspace{1cm}$\text{arrow}\in\{0,1\},\text{dashed}\in\{0,1\})$
  \end{tabular}&\begin{tabular}{l}
    Line from $(x_1,y_1)$ to  $(x_2,y_2)$,\\\hspace{1cm}optionally with an arrow and/or dashed
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|STOP|
  \end{tabular}&\begin{tabular}{l}
    Finishes trace set inference
    \end{tabular}
\\  \bottomrule
\end{tabular}
\end{table}

>>>>

% if you need to pass options to natbib, use, e.g.:
% \PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2018

% ready for submission
\usepackage{nips_2018}

% to compile a preprint version, e.g., for submission to arXiv, add
% add the [preprint] option:
% \usepackage[preprint]{nips_2018}

% to compile a camera-ready version, add the [final] option, e.g.:
% \usepackage[final]{nips_2018}

% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2018}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{dsfont}
\usepackage{verbatimbox}
\usepackage{listings}
\lstset{basicstyle = \small\ttfamily}
\usepackage{graphicx} % more modern
%\usepackage{epsfig} % less modern
%\usepackage{subfig} 
\usepackage{fancyvrb}

@@

@@

\begin{abstract}
  We introduce a model that learns to convert simple hand drawings
  into graphics programs written in a subset of \LaTeX.~The model
  combines techniques from deep learning and program synthesis.  We
  learn a convolutional neural network that proposes plausible drawing
  primitives that explain an image. These drawing primitives are a
  specification (spec) of what the graphics program needs to draw.
  We learn a model that uses program synthesis techniques to
  recover a graphics program from that spec. These programs have
  constructs like variable bindings, iterative loops, or simple kinds
  of conditionals. With a graphics program in hand, we can correct
  errors made by the deep network and extrapolate drawings.  %% Taken
  %% together these results are a step towards agents that induce useful,
  %% human-readable programs from perceptual input.
\end{abstract}

@@

Human vision is rich -- we infer 3D shape, objects, parts of objects,
and relations between objects -- but vision is also abstract:
we can perceive the radial symmetry of a spiral staircase,
the iterated repetition in the Ising model,
see the forest for the trees, and also the recursion within the trees.
How could we build an agent with similar visual inference abilities?
Here we
cast this problem as program learning,
and take as our goal to learn high--level
graphics programs from images.
The graphics programs we consider draw simple figures like those found in machine learning papers
(Fig.~\ref{firstPageExamples}),
and capture high-level features of a drawing like
symmetry, repetition, and reuse of structure.

%% Vision as a kind of ``inverse graphics'' is an idea that is both intuitive and old,
%% with many diverse recent takes on this framing -- neural network approaches like Attend--Infer--Repeat or Capsule Nets,
%% or more structured approaches leveraging probabilistic programming~\cite{}.
%% This vision--as--inverse--graphics metaphor addresses the important problem of
%% inferring, from an image, the identities and poses of
%% objects. But visual perception is richer than object identity and pose:

 %% How can an agent convert noisy, high-dimensional perceptual input
 %% to a symbolic, abstract object, such as a computer program?  Here we
 %% consider this problem within a graphics program synthesis domain.  We
 %% develop an approach for converting  hand
 %% drawings into executable source code for drawing the original image.
 %% The graphics programs in our domain draw simple figures like those found in
 %% machine learning papers
% (see Fig.~\ref{firstPageExamples}a)
\begin{myverbbox}[\small]{\firstFirstPageCode}
for (i < 3)
 rectangle(3*i,-2*i+4,
           3*i+2,6)
 for (j < i + 1)
  circle(3*i+1,-2*j+5)
\end{myverbbox}
\begin{myverbbox}[\small]{\secondFirstPageCode}
reflect(y=8)
 for(i<3)
  if(i>0)
   rectangle(3*i-1,2,3*i,3)
  circle(3*i+1,3*i+1)
\end{myverbbox}

 \begin{figure}[H]%\vspace{-0.7cm}
  \begin{minipage}[b]{0.35\linewidth}  \centering
\begin{tabular}{ll}
  \includegraphics[width = \exampleImageSize]{figures/expert-60.png}&
%  \includegraphics[width = \exampleImageSize]{figures/expert-5.png}&
%    \includegraphics[width = \exampleImageSize]{figures/expert-17.png}&
    \includegraphics[width = \exampleImageSize]{figures/expert-58.png}\\
  \includegraphics[width = \exampleImageSize]{figures/60.png}&
%  \includegraphics[width = \exampleImageSize]{figures/5.png}&
%    \includegraphics[width = \exampleImageSize]{figures/17.png}&
    \includegraphics[width = \exampleImageSize]{figures/58.png}
\end{tabular}
\subcaption{}
  \end{minipage}
  %
  \begin{minipage}[b]{0.6\linewidth}\centering
    \begin{tikzpicture}
      \node(picture1) at (0,-1) {\includegraphics[height = 1.5cm]{figures/expert-31.png}};
      \node(picture2) [below = 0.3cm of picture1] {\includegraphics[height = 1.5cm]{figures/expert-72-trim.png}};
      \node[right=1cm of picture1] (c1) {\fbox{\firstFirstPageCode}};
      \node[right=1cm of picture2] (c2) {\fbox{\secondFirstPageCode}};
      \draw[very thick,->] (picture1.east)  -- (c1.west);
      \draw[very thick,->] (picture2.east)  -- (c2.west);
    \end{tikzpicture}
    \subcaption{}
  \end{minipage}
  \caption{(a): Model learns to convert hand drawings (top) into \LaTeX~(rendered below). (b) Learns to synthesize high-level graphics program from hand drawing.}\label{firstPageExamples}
\end{figure}

 The key observation behind our work is that inferring the program from an image involves two distinct steps that require different technical approaches. The first step involves inferring what objects make up an image -- for diagrams, these are things like as rectangles, lines and arrows. The second step involves identifying the high-level structure in how the objects were drawn. In Fig. 1(b), it means identifying a pattern in how the circles and rectangles are being drawn that is best described with two nested loops, and which can easily be extrapolated to a bigger diagram.

 Framed as probabilistic inference,
 our two-step factoring cashes out as a generative model where a latent program
 is executed to produce a set of drawing commands,
 which are then rendered to form an image (Fig.~\ref{roadmap}).
 We refer to this set of drawing commands as a \textbf{specification (spec)} because it specifies what the graphics program drew while lacking the high-level structure determining how the program decided to draw it.
 We infer the spec from an image using stochastic search (Sequential Monte Carlo)
 and infer a program from a spec using  \textbf{constraint-based program synthesis} \citep{solar2008program} --
 synthesizing structures like symmetries, loops, or conditionals.
 In practice, both stochastic search and program synthesis are
 prohibitively slow,
 and so we learn models that accelerate inference for both programs and specs,
 in the spirit of ``amortized inference''~\cite{paige2016inference},
 using a neural network to amortize the cost of inferring specs from images and using
 a variant of Adaptive Levin Search~\cite{schmidhuber1997shifting}
 to amortize the cost of synthesizing programs from specs.
 \begin{figure}[h]
  \centering\begin{tikzpicture}
    \node[ thick,anchor = west,inner sep=0pt,label={[yshift = 0.3cm]{\small Image}}](observation) at (0,0) {\includegraphics[width = 1.5cm]{figures/expert-39-trimmed.png}};
    \node[ultra thick,anchor = west,inner sep=0pt](traceSource) at (3,0.5){    \begin{lstlisting}[basicstyle = \scriptsize\ttfamily]
line, line,
rectangle,
line, ...
\end{lstlisting}};
    \node[ultra thick,anchor = west,inner sep=0pt](traceImage) at (3.3,-0.5) {
      \includegraphics[width = 0.9cm]{figures/39-parse.png}}; 
    \node(trace)[draw,thin,fit = (traceImage) (traceSource), label = above:{{\small Spec}}] {};

    \node[draw, thick,anchor = west,inner sep=2pt,label=above:{\small Program}](program) at (6.3,0) {
      \begin{lstlisting}
for (j < 3)
for (i < 3)
if (...)
 line(...)
 line(...)
rectangle(...)
    \end{lstlisting}};
    \node[ultra thick,anchor = west,inner sep=0pt,label=below:{\small Extrapolation}](extrapolate) at (9.9,0.5) {\includegraphics[width = 1.3cm]{figures/39-extrapolated.png}};
%    \node[ultra thick,anchor = west,inner sep=0pt,label=below:{\small Similarity}](similarity) at (11.3,0) {\includegraphics[width = 1cm]{figures/expert-38-trim.png}};
    \node[ultra thick,anchor = west,text width = 1.3cm,inner sep=0pt](errors) at (10,-1.2) {\small Error correction};

    \draw[->, ultra thick] (trace.west) -- node[fill = white,rotate = -90] {{\small Rendering}} (observation.east);
    \draw[->, thick, red] ([xshift = -5]observation.south) to[out = -30,in = -150] node[midway,yshift = -6]{{\small Learning + Stochastic search}} ([xshift = -15]trace.south);
    \draw[->, ultra thick] (program.west) -- node[fill = white,rotate = -90] {{\small Execution}} (trace.east);
    \draw[->, thick, red] ([xshift = 10]trace.south) to[out = -10,in = -170] node[midway,yshift = -6]{{\small Learning + Program synthesis}} ([xshift = 10]program.south);
%    \draw[->, thick, red] (program.east) to[out = 80,in = 180] (errors.west);
%    \draw[->, thick] (program.east) to[out = 40,in = -230] (similarity.west);
    \draw[->, thick, red] (program.east) to[out = 60,in = 180] (extrapolate.west);

    \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
    ([yshift = -15,xshift = -115]trace.south) -- node[below = 6pt] {{\small Section~\ref{neuralNetworkSection}: Image$\to$Spec}} ([yshift = -15,xshift = -5]trace.south);
    \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
    ([xshift = 5,yshift = -15]trace.south) -- node[below = 6pt] {{\small Section~\ref{programSynthesisSection}: Spec$\to$Program}} ([xshift = 140,yshift = -15]trace.south);
    \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
([xshift = 150,yshift = -15]trace.south) -- node[below = 6pt] {{\small Section~\ref{applicationsSection}: Applications}} ([xshift = 230,yshift = -15]trace.south);
  \end{tikzpicture}
  \caption{Black arrows: Top--down generative model; Program$\to$Spec$\to$Image. Red arrows: Bottom--up inference procedure.}\label{roadmap}
\end{figure}

\section{Neural architecture for inferring specs}\label{neuralNetworkSection}

We developed a deep network architecture for efficiently inferring a
spec, $S$, from a hand-drawn image, $I$.
Our model combines ideas from
Neurally-Guided Procedural Models~\citep{ritchie2016neurally}
and Attend-Infer-Repeat~\citep{eslami1603attend}, but
we wish to emphasize
that one could use
many different approaches from the computer vision toolkit to
parse an image into primitive drawing commands  (in our terminology, a ``spec'')~\cite{nsd}.
Our network constructs the
spec one drawing command at a time, conditioned on what it has drawn so far (Fig.~\ref{architecture}).
%When predicting the next drawing
%command, the network takes as input the target image $I$ as well as
%the rendered output of previous drawing commands.
%Intuitively, the network looks at the image it wants to explain, as well as what it has already drawn.  It then decides either to stop drawing or proposes another drawing command to add to the trace set; if it decides to continue drawing, the predicted primitive is rendered to its ``canvas'' and the process repeats.
We first
pass a $256\times 256$ target image and a rendering of the drawing commands so
far (encoded as a two-channel image) to a convolutional network. Given
the features extracted by the convnet, a multilayer perceptron then
predicts a distribution over the next drawing command to execute
(see Tbl.~\ref{drawingCommandTable}).
%We predict the drawing command token-by-token,
%\footnote{Tokens are atomic symbols in the programming language  -- so the model is \emph{not} predicting lines of code character-by-character, but instead the model knows about the syntactic structure of primitive drawing commands}
%%   conditioning each token both on the image
%% features and on the previously generated tokens. 
We also use a
differentiable attention mechanism (Spatial Transformer
Networks:~\cite{jaderberg2015spatial}) to let the model attend to
different regions of the image while predicting drawing commands.
%In Fig.~\ref{architecture} the network first decided to emit the \verb|circle| token
%conditioned on the image features, then it emitted the $x$ coordinate of
%the circle conditioned on the image features and the \verb|circle|
%token, and finally it predicted the $y$ coordinate of the circle
%conditioned on the image features, the \verb|circle| token, and the
%$x$ coordinate.
We currently constrain
coordinates to lie on a discrete $16\times 16$ grid,
but the grid could be made arbitrarily fine.  %% See supplement for the full details
%% of the architecture, which we implemented in
%% Tensorflow~\citep{tensorflow2015-whitepaper}.
%% \remark{There are some more details that are important to provide
%%   about this architecture in the supplement: the
%%   functional form(s) of the probability distributions over tokens, the
%%   network layer sizes, which MLPs share parameters, etc.}

@@

  \node[draw,ultra thick,anchor = west,inner sep=0pt,minimum width = 2cm,minimum height = 3cm] (CNN) at (4.5,-2.5) {CNN};
  \node[inner sep = 0pt](tensorProduct) at ([xshift = -1.5cm]CNN.west) {$\bigoplus$};

@@

%  \node at (canvas.x,canvas.y) {Canvas};
\end{tikzpicture}
\caption{Our neural architecture for inferring the spec of a graphics program from its output. \textcolor{blue}{Blue}: network inputs. Black: network operations. \textcolor{red}{Red}: samples from a multinomial. \texttt{Typewriter font}: network outputs. Renders snapped to a $16\times 16$ grid, illustrated in \textcolor{gray}{gray}. STN (spatial transformer network) is a differentiable attention mechanism~\citep{jaderberg2015spatial}.}  \label{architecture}
\end{figure}
\begin{table}[t]
  \caption{Primitive drawing commands currently supported by our model.}
\label{drawingCommandTable}
\begin{tabular}{ll}\toprule
  \begin{tabular}{l}
    \verb|circle|$(x,y)$
  \end{tabular}& \begin{tabular}{l}
    Circle at $(x,y)$
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|rectangle|$(x_1,y_1,x_2,y_2)$
  \end{tabular}&\begin{tabular}{l}
    Rectangle with corners at $(x_1,y_1)$ \& $(x_2,y_2)$
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|line|$(x_1,y_1,x_2,y_2,$\\
    \hspace{1cm}$\text{arrow}\in\{0,1\},\text{dashed}\in\{0,1\})$
  \end{tabular}&\begin{tabular}{l}
    Line from $(x_1,y_1)$ to  $(x_2,y_2)$,\\\hspace{1cm}optionally with an arrow and/or dashed
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|STOP|
  \end{tabular}&\begin{tabular}{l}
    Finishes trace set inference
    \end{tabular}
\\  \bottomrule
\end{tabular}
\end{table}

%% For the model in Fig.~\ref{architecture}, the distribution over the next drawing command factorizes as:
%% \begin{equation}
%%   \probability_\theta [t_1t_2\cdots t_K | I,T] = \prod_{k = 1}^K \probability_\theta \left[t_k | a_\theta \left(f_\theta(I,\text{render}(T)) | \{t_j\}_{j = 1}^{k - 1}\right), \{t_j\}_{j = 1}^{k - 1}\right]
%% \end{equation}
%% where $t_1t_2\cdots t_K$ are the tokens in the drawing command, $I$ is
%% the target image, $T$ is a trace set, $\theta$ are the
%% parameters of the neural network, $f_\theta(\cdot,\cdot)$ is the
%% image feature extractor (convolutional network), and $a_\theta(\cdot|\cdot)$ is an attention mechanism. The distribution over
%% traces factorizes as:
%% \begin{equation}
%%   \probability_\theta [T|I] = \prod_{n = 1}^{|T|} \probability_\theta [T_n | I,T_{1:(n-1)}]\times\probability_\theta [\verb|STOP| | I,T]\label{objective}
%% \end{equation}
%% where $|T|$ is the length of trace set $T$, the subscripts
%% on $T$ index drawing commands within the trace (so $T_n$ is a sequence of tokens: $t_1t_2\cdots t_K$), and the \verb|STOP|
%% token is emitted by the network to signal that the trace set
%% explains the image.
We trained our network by sampling specs $S$ and target
images $I$ for randomly generated scenes
and maximizing
%Eq.~\ref{objective}
 the likelihood of $S$ given $I$ with respect to
 the model parameters, $\theta$, by gradient ascent.
We trained on $10^5$ scenes, which takes a day on an Nvidia TitanX GPU.
